---
title: "DREF Forecasting"
author: "Hamish Patten"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(magrittr)
library(tidyverse)
library(ggplot2)

GetMonty_ID<-function(DF,haz=NULL) {
  # In case a specific hazard is fed in
  if(!is.null(haz)) DF%<>%mutate(haz_Ab=haz)
  # Generate the names from the dataframe
  namerz<-DF%>%
    dplyr::select(haz_Ab,ev_sdate,ev_ISO3s)%>%
    apply(1,function(x) paste0(x,collapse = "-"))
  
  paste0(namerz,"-GCDB_V1")
}

GetGCDB_impID<-function(impies){
  # For empty spatial IDs
  if(any(colnames(impies)=="imp_spat_ID")) impies$tmp<-unlist(lapply(impies$imp_spat_ID, function(x) paste0(x,collapse=":")))
  # Compile the ID
  impies$imp_sub_ID<-impies%>%dplyr::select(dplyr::any_of(c("event_ID","imp_src_db","exp_spec","imp_type","imp_sdate","imp_fdate","imp_units","tmp","imp_lon","imp_lat")))%>%
    mutate(imp_src_db=stringr::str_remove(stringi::stri_trans_totitle(imp_src_db),pattern = " "))%>%
    apply(1,function(x) paste0(x,collapse = "-"))
  # Make sure that no duplicates exist...
  impies%>%group_by(imp_sub_ID)%>%
    mutate(imp_sub_ID=paste0(imp_sub_ID,"_",1:length(imp_sub_ID)))%>%
    dplyr::select(-any_of(c("tmp")))%>%as.data.frame()
}

convIso3Continent_alt<-function(iso3){
  # continents<-countrycode::countrycode(sourcevar = iso3,
  #                                      origin = "iso3c",
  #                                      destination = "continent",warn = F)
  left_join(data.frame(ISO3=iso3),raster::ccodes()[,c("ISO3","continent")],by="ISO3")$continent
}

AsYear<-function(date,red=F,limit=T){
  str_split(as.character(date),pattern = "-",simplify = T)[,1]%>%as.integer()%>%return()
}

AsMonth<-function(date){
  str_split(as.character(date),pattern = "-",simplify = T)[,2]%>%as.integer()%>%return()
  # return(as.numeric(format(as.Date(date),"%m")))
}

getGOurl<-function(db="GO-App",token=NULL, results=T){
  
  if(db=="GO-App") {db_code<-"appeal"
  } else if(db=="GO-FR") {db_code<-"field_report"
  } else if(db=="GO-DREF") {db_code<-"dref-final-report"
  } else stop("this IFRC-GO database does not exist")
  
  # Base URL
  url<-paste0("https://goadmin.ifrc.org/api/v2/",db_code,"/?format=json&limit=100000000000") 
  # It's easier if we only want to access the public-facing data...
  if(!is.null(token)){
    # Make a request using a token
    req <- httr::GET(url, httr::add_headers(Authorization = token),httr::timeout(10000))
    # Extract the content
    json <- httr::content(req, as = "text", encoding = "UTF-8")
    # Convert from JSON to R list or tibble output
    json%<>%jsonlite::fromJSON()
  } else json<-jsonlite::fromJSON(url)
  
  if(results) return(json$results) else return(json)
}

ImpLabs<-function(ImpDB,nomDB="Desinventar",dropName=T){
  # Open up the database impact taxonomy conversion file
  imptax<-openxlsx::read.xlsx("../../Taxonomies/ConvertFromDatabases/ConvertImpact_Taxonomy.xlsx")%>%
    filter(imp_src_db==nomDB)
  # Find where the Desinventar data impact estimates stop 
  vlim<-which(colnames(ImpDB)%in%imptax$VarName)
  # For all columns that correspond to impact estimates, return the data
  ImpDB%<>%reshape2::melt(measure.vars=colnames(ImpDB)[vlim])%>%
    mutate(VarName=as.character(variable),imp_value=as.numeric(value))%>%
    dplyr::select(-c(variable,value))%>%
    left_join(dplyr::select(imptax,-c("imp_src_orgtype","imp_src_org","imp_src_db")),by="VarName")
  # Spit it out!
  if(dropName) return(ImpDB) else return(dplyr::select(ImpDB,-VarName))
}

ExtractGOdata<-function(db="GO-App", token = NULL, results=T){
  options(timeout = max(10000, getOption("timeout")))
  getGOurl(db=db,token,results)
}

DREF<-ExtractGOdata(db = "GO-App")

CleanGO_app<-function(appeal){
  
  appeal$haz_Ab<-appeal$haz_spec<-appeal$dtype$name
  appeal$dtype<-NULL
  
  appeal$imp_ISO3s<-appeal$country$iso3
  appeal$region<-convIso3Continent_alt(appeal$imp_ISO3s)

  appeal$country<-NULL
  
  appeal$start_date<-str_split(appeal$start_date,"T",simplify = T)[,1]
  appeal$end_date<-str_split(appeal$end_date,"T",simplify = T)[,1]
  
  appeal%<>%mutate(imp_ISO3s=imp_ISO3s,ev_ISO3s=imp_ISO3s,region=region,
                   ev_name=str_replace_all(name,'"',""),location=str_replace_all(name,'"',""),
                   ev_name_lang="lang_eng",
                   imp_sdate=as.character(as.Date(start_date)),imp_fdate=as.character(as.Date(end_date)),
                   ev_sdate=as.character(as.Date(start_date)),ev_fdate=as.character(as.Date(end_date)),
                   # ev_sdate=as.character(as.Date(created_at)),ev_fdate=as.character(as.Date(created_at)),
                   imp_unitdate=as.character(as.Date(modified_at)),
                   imp_src_URL="https://goadmin.ifrc.org/api/v2/appeal",
                   imp_src_orglab="International Federation of Red Cross and Red Crescent Societies (IFRC)",
                   imp_src_org="IFRC",
                   imp_src_db=atype_display,
                   imp_src_orgtype="orgtypengo",
                   imp_spat_covcode="spat_polygon",
                   imp_spat_res=0,
                   imp_spat_resunits="adminlevel",
                   imp_spat_crs="EPSG:4326",
                   imp_spat_srcorg="IFRC",
                   imp_spat_srcdb="GO-Maps",
                   imp_spat_URL="https://go-user-library.ifrc.org/maps",
                   imp_spat_ID=NA_character_)
  
  appeal$event_ID<-GetMonty_ID(appeal,haz=appeal$haz_Ab)
  
  appeal%<>%ImpLabs(nomDB = "GO-App", dropName = T)
  
  # Create an impact-specific ID
  appeal%<>%GetGCDB_impID()
  # Add missing columns & reorder the dataframe to fit imp_GCDB object
  # appeal%<>%AddEmptyColImp()
  
  # appeal$GLIDE<-GetGLIDEnum(appeal)
  
  # Make sure to remove repeated entries for when they update them
  appeal%<>%mutate(aid=as.integer(aid))%>%arrange(desc(aid))%>%
    filter(!duplicated(imp_sub_ID))%>%arrange(aid)
  
  counties<-openxlsx::read.xlsx("../../Taxonomies/IsoContinentRegion.xlsx")%>%
    filter(!is.na(Country))%>%mutate(Continent=convIso3Continent_alt(ISO.Code))%>%
    dplyr::select(ISO.Code,Country,UN.Region,World.Bank.Regions,Continent,UN.Sub.Region,World.Bank.Income.Groups)%>%
  setNames(c("ISO3","country","unregion","worldbankregion","continent","unsubregion","worldbankincomegroup"))
  
  appeal%<>%left_join(counties,by=c("ev_ISO3s"="ISO3"))
  
  appeal%<>%filter(imp_src_db=="DREF")
  
  # Also group hazards to make up the numbers
  appeal$haz_Ab_grp<-"Non-CC"
  appeal$haz_Ab_grp[appeal$haz_Ab%in%c("Cold Wave","Cyclone","Drought","Fire","Flood","Heat Wave","Pluvial/Flash Flood","Storm Surge")]<-"CC"
  # Extract the year information
  appeal$year<-AsYear(appeal$ev_sdate)

  appeal%<>%filter(ev_sdate>=as.Date("2008-01-01") & imp_src_db=="DREF")
  
  return(appeal)
}

# # Get the DREF data and convert into tabular format
tDREF<-CleanGO_app(DREF); rm(DREF)

tDREF$haz_Ab[tDREF$haz_Ab%in%c("Fire","Drought","Heat Wave")]<-"Heat-Related"
tDREF$haz_Ab[tDREF$haz_Ab%in%c("Pluvial/Flash Flood")]<-"Flood"
# tEMDAT<-convEMDAT_Monty()%>%procDBfore

```

## Exploratory Data Analysis

Let's first have a look at some basic analysis of the data. In order to accurately predict the DREF funding, we need to understand the number of events that had any funding allocated and the amount of funding allocated per event, per year. We want to look at spatio-temporal trends on two levels: seasonal per continent/region and yearly. Furthermore, the allocation per hazard is really important.

```{r cars}
# Check the average number of events + total yearly funding allocated
tDREF%>%mutate(ev_sdate=as.Date(ev_sdate))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall")%>%
  group_by(year)%>%reframe(Count=length(year),DREF=median(imp_value))%>%
  ggplot()+geom_point(aes(year,Count,size=DREF,colour=DREF))+geom_line(aes(year,Count))
# Total monthly average funding, since 2021
tDREF%>%mutate(ev_sdate=as.Date(ev_sdate),
               month=as.factor(month(ev_sdate)))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall")%>%
  group_by(year,month)%>%reframe(Count=length(year),DREF=sum(imp_value))%>%
  ggplot()+geom_boxplot(aes(month,DREF,fill=month))#+geom_line(aes(month,DREF))
# Check the number of past two years monthly funding allocations
tDREF%>%mutate(ev_sdate=as.Date(ev_sdate),
               month=(as.integer(month(ev_sdate))/12-0.5/12))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall" & year>2021)%>%
  group_by(year,month)%>%reframe(Count=length(year),DREF=sum(imp_value),yrmonth=year+month)%>%
  ggplot()+geom_point(aes(yrmonth,Count,size=DREF,colour=DREF))+geom_line(aes(yrmonth,Count))
# Check the past two years monthly funding allocations
tDREF%>%mutate(ev_sdate=as.Date(ev_sdate),
               month=(as.integer(month(ev_sdate))/12-0.5/12))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall" & year>2021)%>%
  group_by(year,month)%>%reframe(Count=length(year),DREF=sum(imp_value),yrmonth=year+month)%>%
  ggplot()+geom_point(aes(yrmonth,DREF,colour=Count))+geom_line(aes(yrmonth,DREF))

# Check the average number of events + total monthly funding allocated
tDREF%>%mutate(ev_sdate=as.Date(ev_sdate))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall")%>%
  group_by(year)%>%reframe(Count=length(year),DREF=median(imp_value))%>%
  ggplot()+geom_point(aes(year,Count,size=DREF,colour=DREF))+geom_line(aes(year,Count))

# Check only the median yearly funding allocated per event
tDREF%>%mutate(ev_sdate=as.Date(ev_sdate))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall")%>%
  group_by(year)%>%reframe(Count=length(year),
                                     Sum=sum(imp_value),
                                     Cost=median(imp_value))%>%
  ggplot()+geom_point(aes(year,Cost,size=Count,colour=Sum))+
  geom_text(aes(year,Cost+2e4,label=Count))+
  geom_line(aes(year,Cost))+
  xlab("Year")+ylab("Yearly Median Funding Allocated Per Event")
# Split by region
tDREF%>%mutate(ev_sdate=as.Date(ev_sdate))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall")%>%
  group_by(year,worldbankregion)%>%reframe(Count=length(year),
                           Sum=sum(imp_value),
                           Cost=median(imp_value))%>%
  ggplot()+geom_point(aes(year,Cost,size=Count,colour=Sum))+
  geom_text(aes(year,Cost+10e4,label=Count))+
  geom_line(aes(year,Cost))+
  xlab("Year")+ylab("Yearly Median Funding Allocated Per Event")+
  facet_wrap(~worldbankregion,ncol = 2)
# Split by hazard
tDREF$Hazard<-tDREF$haz_Ab_grp
tDREF$Hazard[tDREF$haz_Ab%in%c("Pluvial/Flash Flood","Flood")]<-"Flood"

tDREF%>%mutate(ev_sdate=as.Date(ev_sdate))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall")%>%
  group_by(year,Hazard)%>%reframe(Count=length(year),
                           Sum=sum(imp_value),
                           Cost=median(imp_value))%>%
  ggplot()+geom_point(aes(year,Cost,size=Count,colour=Sum))+
  geom_text(aes(year,Cost+10e4,label=Count))+
  geom_line(aes(year,Cost))+
  xlab("Year")+ylab("Yearly Median Funding Allocated Per Event")+
  facet_wrap(~Hazard,ncol = 2)
# Median funding allocated by month
tDREF%>%mutate(ev_sdate=as.Date(ev_sdate),
               month=as.integer(month(ev_sdate)))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall" & haz_Ab_grp=="CC")%>%
  group_by(month,worldbankregion)%>%reframe(Count=length(year),
                           Sum=sum(imp_value),
                           Cost=median(imp_value))%>%
  ggplot()+geom_point(aes(month,Cost,size=Count,colour=Sum))+
  # geom_text(aes(month,Cost+10e4,label=Count))+
  geom_line(aes(month,Cost))+
   scale_x_continuous(breaks=1:12,labels=month.abb)+
  xlab("Month")+ylab("Monthly Median Funding Allocated Per Event")+
  facet_wrap(~worldbankregion,ncol = 2)
# Number of events per month
tDREF%>%mutate(ev_sdate=as.Date(ev_sdate),
               month=as.integer(month(ev_sdate)))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall" & haz_Ab_grp=="CC")%>%
  group_by(month,worldbankregion)%>%reframe(Count=length(year),
                           Sum=sum(imp_value),
                           Cost=median(imp_value))%>%
  ggplot()+geom_point(aes(month,Count,size=Cost,colour=Cost))+
  # geom_text(aes(month,Cost+10e4,label=Count))+
  geom_line(aes(month,Count))+
   scale_x_continuous(breaks=1:12,labels=month.abb)+
  ylim(c(0,NA))+
  xlab("Month")+ylab("Number of Events")+
  facet_wrap(~worldbankregion,ncol = 2,scales = "free")
# Look at monthly seasonality of different hazards in sub-saharan africa
tDREF%>%mutate(ev_sdate=as.Date(ev_sdate),
               month=as.integer(month(ev_sdate)))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall" & 
           worldbankregion=="Sub-Saharan Africa")%>%
  group_by(month,Hazard)%>%reframe(Count=length(year),
                           Sum=sum(imp_value),
                           Cost=median(imp_value))%>%
  ggplot()+geom_point(aes(month,Count,size=Cost,colour=Cost))+
  # geom_text(aes(month,Cost+10e4,label=Count))+
  geom_line(aes(month,Count))+
   scale_x_continuous(breaks=1:12,labels=month.abb)+
  ylim(c(0,NA))+
  xlab("Month")+ylab("Number of Events")+
  facet_wrap(~Hazard,ncol = 2,scales = "free")

```

```{r floods,fig.width=10,fig.height=20}
# Now let's look specifically at floods (large amount of statistics) for each unsubregion
tDREF%>%mutate(ev_sdate=as.Date(ev_sdate),
               month=as.integer(month(ev_sdate)))%>%arrange(ev_sdate)%>%
  filter(exp_spec=="expspec_ecoifrcall" & 
           Hazard=="Flood")%>%
  group_by(month,unsubregion)%>%reframe(Count=length(year),
                           Sum=sum(imp_value),
                           Cost=median(imp_value))%>%
  ggplot()+geom_point(aes(month,Count,size=Cost,colour=Cost))+
  # geom_text(aes(month,Cost+10e4,label=Count))+
  geom_line(aes(month,Count))+
   scale_x_continuous(breaks=1:12,labels=month.abb)+
  ylim(c(0,NA))+
  xlab("Month")+ylab("Number of Events")+
  facet_wrap(~unsubregion,ncol = 2,scales = "free")

```

Let's start modelling!

```{r Modelling}

# Set which months to focus on:
dateyyy<-Sys.Date()
monthyyy<-AsMonth(dateyyy)
yearyyy<-AsYear(dateyyy)

# tDREF%>%mutate(ev_sdate=as.Date(ev_sdate),events=1)%>%arrange(ev_sdate)%>%
#   filter(exp_spec=="expspec_ecoifrcall")%>%
#   mutate(rollav=zoo::rollapply(events,150,function(x) sum(x,na.rm = T),align='right',fill=NA))%>%
#   ggplot()+geom_line(aes(ev_sdate,rollav))+
#   scale_y_log10()

predVal<-function(year,imp_value,se=F){
  if(!se) return(predict(lm(imp_value~year,data.frame(imp_value=imp_value,year=year)),data.frame(year=yearyyy)))
  return(predict(lm(imp_value~year,
                    data.frame(imp_value=imp_value,year=year)),
                 data.frame(year=yearyyy),
                 se.fit=T)$se.fit)
}

# Calculate the number of monthly events in 2024 as a regression since 2018
nevs<-tDREF%>%
  filter(exp_spec=="expspec_ecoifrcall" & 
           year>2018 & year<2024)%>%
  group_by(haz_Ab_grp,haz_Ab,year,worldbankregion)%>%
  reframe(Events=length(imp_value))%>%
  group_by(haz_Ab_grp,haz_Ab,worldbankregion)%>%
  reframe(Month=month.abb,
          numMonth=1:12,
          Evs=unname(pmax(predVal(year,Events),median(Events[year>(yearyyy-5)],na.rm = T),na.rm = T)/24),
          lowEvs=pmin(unname(pmax(Evs-predVal(year,Events,se=T)/24,min(Events,na.rm = T),na.rm = T)),Evs),
          uppEvs=pmax(unname(pmin(Evs+predVal(year,Events,se=T)/24,max(Events,na.rm = T),na.rm = T)),Evs))

# Now calculate the 2024 allocated funding as a regression since 2018
expcost<-tDREF%>%
  filter(exp_spec=="expspec_ecoifrcall" & 
           year>2018 & year<2024)%>%
  group_by(haz_Ab_grp,haz_Ab,worldbankregion)%>%
  # reframe(Alloc=median(imp_value))%>%
  # group_by(haz_Ab,worldbankregion)%>%
  reframe(Month=month.abb,
          numMonth=1:12,
          Cst=unname(pmax(predVal(year,imp_value),median(imp_value[year>(yearyyy-5)],na.rm=T),na.rm = T)),
          lowCst=pmin(unname(pmax(Cst-predVal(year,imp_value,se=T),min(imp_value,na.rm = T),na.rm = T)),Cst),
          uppCst=pmax(unname(pmin(Cst+predVal(year,imp_value,se=T),max(imp_value,na.rm = T),na.rm = T)),Cst))
# Ensure any issues with the standard error are resolved by setting confidence intervals to the expected value
# expcost$Cst[is.infinite(expcost$Cst) & is.infinite(expcost$lowCst) & is.infinite(expcost$uppCst) |
#               expcost$Cst<1e-5 & expcost$lowCst<1e-5 & expcost$uppCst<1e-5]<-NA
# # If issues with the 
# expcost$lowCst[is.infinite(expcost$lowCst) & is.infinite(expcost$uppCst)]<-expcost$Cst
# expcost$uppCst[is.infinite(expcost$lowCst) & is.infinite(expcost$uppCst)]<-expcost$Cst
# 
# expcost$lowCst[expcost$lowCst<1e-5 & expcost$Cst==expcost$uppCst]<-
#   expcost$Cst[expcost$lowCst<1e-5 & expcost$Cst==expcost$uppCst]
# Any value where the expected allocated funding is zero are removed
# expcost[expcost$Cst<1e-5,6:8]<-NA
expcost%<>%filter(!is.na(Cst))

nonCCcost<-left_join(nevs,expcost,
                     by = join_by(haz_Ab_grp, haz_Ab, worldbankregion, Month, numMonth))%>%
  mutate(DREF=Evs*Cst,lowDREF=lowEvs*lowCst,uppDREF=uppEvs*uppCst)

# For the climate change related hazards, add in the seasonality
Seasonality<-tDREF%>%filter(haz_Ab_grp=="CC" & exp_spec=="expspec_ecoifrcall" & 
                 !is.na(worldbankregion) & worldbankregion!="Not Classified")%>%
  mutate(Month=as.integer(month(ev_sdate)))%>%
  group_by(haz_Ab_grp,haz_Ab,worldbankregion,Month)%>%
  reframe(Evs=length(year))
  
# Smooth this using a 1-month-either-side rolling average
Seasonality%<>%rbind(Seasonality%>%mutate(Month=Month-12),
                     Seasonality%>%mutate(Month=Month+12))%>%
  arrange(Month)%>%
  group_by(haz_Ab_grp,haz_Ab,worldbankregion)%>%
  mutate(Evs=zoo::rollapply(Evs,3,function(x) mean(x,na.rm = T),align='center',fill=NA))%>%
  filter(Month>=1 & Month<=12)

Seasonality%>%ggplot()+geom_point(aes(Month,Evs,colour=haz_Ab))+
  # geom_text(aes(Month,Cost+10e4,label=Evs))+
  geom_line(aes(Month,Evs,colour=haz_Ab))+
   scale_x_continuous(breaks=1:12,labels=month.abb)+
  ylim(c(0,NA))+
  xlab("Month")+ylab("Number of Events")+
  facet_wrap(~worldbankregion,ncol = 2,scales = "free")

# Scale it
Seasonality%<>%group_by(haz_Ab_grp,haz_Ab,worldbankregion)%>%
  mutate(Month=Month,PDF=Evs/sum(Evs))

# Now combine it with the rest of the non-CC data
CCcost<-nonCCcost%>%filter(numMonth==1 & haz_Ab_grp=="CC")%>%
  dplyr::select(-c(Month,numMonth))%>%
  left_join(Seasonality%>%dplyr::select(-Evs),
            by = join_by(haz_Ab_grp, haz_Ab, worldbankregion))%>%
  mutate(DREF=12*Evs*Cst*PDF,lowDREF=12*lowEvs*lowCst*PDF,uppDREF=12*uppEvs*uppCst*PDF,
         numMonth=Month,Month=month.abb[Month])

foreDREF<-rbind(CCcost%>%dplyr::select(-PDF),
                nonCCcost%>%filter(haz_Ab_grp=="Non-CC"))

# General plot of DREF funding allocation per month in 2024, per region, CC vs non-CC
foreDREF%>%group_by(haz_Ab_grp,worldbankregion,numMonth)%>%
    reframe(DREF=sum(DREF),lowDREF=sum(lowDREF),uppDREF=sum(uppDREF))%>%
    ggplot()+geom_smooth(aes(numMonth,DREF,colour=haz_Ab_grp))+
    scale_x_continuous(breaks=1:12,labels=1:12)+
    facet_wrap(~worldbankregion)

# General plot of DREF funding allocation per month in 2024, per region, CC vs non-CC
foreDREF%>%group_by(haz_Ab,numMonth)%>%
    reframe(DREF=sum(DREF),lowDREF=sum(lowDREF),uppDREF=sum(uppDREF))%>%
    ggplot()+geom_line(aes(numMonth,DREF,colour=haz_Ab))+
  geom_point(aes(numMonth,DREF,colour=haz_Ab))+
    scale_x_continuous(breaks=1:12,labels=1:12)+
    facet_wrap(~haz_Ab)

# Filter for only the months we care about
foreDREF%<>%filter(Month%in%month.abb[month(rep(dateyyy+c(0,30,60)))])%>%
  arrange(numMonth)%>%rename(c("Hazard"="haz_Ab","Region"="worldbankregion"))

filename <- paste0("../../Analysis_Results/DREF_forecast_",Sys.Date(),".xlsx")
# Create the xlsx workbook
wb<-openxlsx::createWorkbook()
# Create all the worksheets
wb%>%openxlsx::addWorksheet("Total");wb%>%openxlsx::addWorksheet("Per Region");wb%>%openxlsx::addWorksheet("Per Hazard");wb%>%openxlsx::addWorksheet("Full Dataset")
# Events
openxlsx::writeData(wb, 
                    sheet="Total",
                    foreDREF%>%group_by(numMonth,Month)%>%
                      reframe(DREF=sum(DREF),lowDREF=sum(lowDREF),uppDREF=sum(uppDREF))%>%
                      dplyr::select(-numMonth), 
                    headerStyle=openxlsx::createStyle(textDecoration = "Bold"),
                    keepNA = F)
openxlsx::writeData(wb, 
                    sheet="Per Region",
                    foreDREF%>%group_by(numMonth,Month,Region)%>%
                      reframe(DREF=sum(DREF),lowDREF=sum(lowDREF),uppDREF=sum(uppDREF))%>%
                      group_by(numMonth,Month)%>%
                      mutate(Percentage=100*DREF/sum(DREF))%>%ungroup()%>%
                      dplyr::select(-numMonth),
                    headerStyle=openxlsx::createStyle(textDecoration = "Bold"),
                    keepNA = F)
openxlsx::writeData(wb, 
                    sheet="Per Hazard",
                    foreDREF%>%group_by(numMonth,Month,Hazard)%>%
                      reframe(Month=unique(Month),DREF=sum(DREF),lowDREF=sum(lowDREF),uppDREF=sum(uppDREF))%>%
                      group_by(numMonth,Month)%>%
                      mutate(Percentage=100*DREF/sum(DREF))%>%ungroup()%>%
                      dplyr::select(-numMonth), 
                    headerStyle=openxlsx::createStyle(textDecoration = "Bold"),
                    keepNA = F)
openxlsx::writeData(wb, 
                    sheet="Full Dataset",
                    foreDREF%>%group_by(numMonth,Month,Region,Hazard)%>%
                      reframe(Month=unique(Month),DREF=sum(DREF),lowDREF=sum(lowDREF),uppDREF=sum(uppDREF))%>%
                      group_by(numMonth,Month)%>%
                      mutate(Percentage=100*DREF/sum(DREF))%>%ungroup()%>%
                      dplyr::select(-numMonth), 
                    headerStyle=openxlsx::createStyle(textDecoration = "Bold"),
                    keepNA = F)

# Save it out!
openxlsx::saveWorkbook(wb,filename,overwrite = T)


```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
