---
title: Monty Pairing
subtitle: "Hamish Patten, Karla Ayivi, Marcela Duran Arias, <br> Arun Gandhi and Jemimah Ndugwa"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    css: styles.css
---

<style type="text/css">
h1.title {
  font-size: 50px;
  color: White;
}
h3.subtitle {
  font-size: 20px;
  color: White;
}
<!-- .table-hover > tbody > tr:hover { -->
<!--   background-color: #f7f7e6; -->
<!-- } -->
.main-container {
  width: 95%;
    <!-- max-width: unset; -->
  }
</style>

<div style= "position:relative">
<!-- <div style= "float:left; position:relative; top:-100px;margin-bottom:-100px;margin-left:-10px;"> -->

```{r setup, include=FALSE}
library(dplyr)
library(magrittr)
library(tidyverse)
library(kableExtra)
library(knitr)
library(prettydoc)
library(rmdformats)
library(caret)
library(parallel)
library(doParallel)

knitr::opts_chunk$set(echo = TRUE)
options(kableExtra.latex.load_packages = FALSE,
        knitr.kable.NA = '')
# Read in the (modified) data
pairies<-readRDS("../../Analysis_Results/Pairing/PairedData.RData")
# Labels for the hazard codes
haz_Ab_lab<-c("AV"="Avalanche",
              "CW"="Cold Wave",
              "DR"="Drought",
              "EQ"="Earthquake",
              "EP"="Epidemic",
              "ER"="Erosion",
              "EC"="Extra-Tropical Cyclone",
              "ET"="Extreme Temperature",
              "FR"="Fire",
              "FF"="Flash Flood",
              "FL"="Flood",
              "HW"="Heat Wave",
              "HT"="High Temperature",
              "IN"="Insect Infestation",
              "LS"="Landslide",
              "MM"="Mass Movement",
              "MS"="Mudslide",
              "SL"="Slide",
              "SN"="Snowfall",
              "ST"="Storm",
              "SS"="Storm Surge",
              "TO"="Tornado",
              "TC"="Tropical Cyclone",
              "TS"="Tsunami",
              "WF"="Wildfire",
              "VW"="Violent Wind",
              "VO"="Volcanic Activity",
              "WV"="Wave",
              "WA"="Wave Action")
# Convert ISO3Cs to continent
convIso3Continent<-function(iso3){
  continents<-countrycode::countrycode(sourcevar = iso3,
                                       origin = "iso3c",
                                       destination = "continent",warn = F)
}
```

</div>

## Database

These are the databases we're working with:

```{r dbs, message=FALSE,warning=FALSE, echo=FALSE}

table(c(pairies$targ_db,pairies$dest_db))%>%as.data.frame()%>%
  mutate(Var2=str_split(Var1," - ",simplify = T)[,1],
         Var1=str_split(Var1," - ",simplify = T)[,2])%>%
  dplyr::select(Var1,Var2,Freq)%>%arrange(desc(Freq))%>%
kable(col.names = c("Organisation","Database","Number of Records Sampled"))

```

<br><br>

The number of events per year:

```{r yrevs, message=FALSE,warning=FALSE, echo=FALSE}

pairies%>%ggplot()+geom_density(aes(as.integer(str_split(targ_evsdate,"-",simplify = T)[,1])),
                                stat = "count")+
  xlim(c(1980,2024))+xlab("Year")+ylab("Number of Records")

```

<br><br>

The number of records per region

```{r isos, message=FALSE,warning=FALSE, echo=FALSE}

table(convIso3Continent(c(pairies$targ_evISOs,pairies$dest_evISOs)))%>%as.data.frame()%>%
  arrange(desc(Freq))%>%
kable(col.names = c("Region","Number of Records Sampled"))

```

<br><br>

We're working with multiple hazards

```{r hazzies, message=FALSE,warning=FALSE, echo=FALSE}
hazzies<-unique(as.vector(str_split(c(pairies$targ_hzAb,pairies$dest_hzAb)," : ",simplify = T)))
hazzies<-hazzies[nchar(hazzies)==2]

tmp<-do.call(rbind,lapply(hazzies,function(haz){
  data.frame(hazAb=haz,Hazard=haz_Ab_lab[haz],
             Count=sum(grepl(haz,c(pairies$targ_hzAb,pairies$dest_hzAb),ignore.case = F)))
}))%>%arrange(desc(Count))

manyhaz<-tmp$hazAb[tmp$Count>10]

tmp%>%kable(col.names = c("Hazard Code","Hazard","Number of Records Sampled"))

```

<br><br>

Here's a sample of the data:

```{r database, message=FALSE,warning=FALSE, echo=FALSE}
pairies%>%
  slice(sample(1:nrow(pairies),20,replace = F))%>%
  dplyr::select(-targ_mid,-targ_evID,-targ_URL,-targ_ID,
                -dest_mid,-dest_evID,-dest_URL,-dest_ID,
                -targ_sday, -targ_fday, -dest_sday, -dest_fday, 
                -paired, -confidence)%>%
  kable(align="l", format = "html", 
        col.names=c("Database","Start Date","End Date","Longitude","Latitude",
                    "Hazard Code","ISO3C",
                    "Database","Start Date","End Date","Longitude","Latitude",
                    "Hazard Code","ISO3C","Distance","Pairing Decision"))%>%
  add_header_above(c("TARGET" = 7, "DESTINATION" = 7, "PAIRING" = 2))%>%
  kable_styling(c("bordered", "condensed"), full_width = T)
```


## Analysis of Manual Pairings

Pairing allocation, per database

```{r confpair, echo=FALSE,message=FALSE,warning=FALSE}
nomies<-colnames(pairies[,c(1:11,23:ncol(pairies))])
tmp<-rbind(pairies[,c(1:11,23:ncol(pairies))],
           pairies[,c(12:ncol(pairies))]%>%setNames(nomies))

ploty<-tmp %>%
  group_by(targ_db, confpair) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count))%>%
  ggplot()+geom_bar(aes(targ_db,percentage,fill=confpair),stat="identity")+
  xlab("Database")+ylab("Percentage of Records")+labs(fill="Pairing")+
  scale_y_continuous(labels = scales::percent_format())+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plotly::ggplotly(ploty)%>% plotly::layout(yaxis = list(hoverformat = '.4f'))
```

<br><br>

Let's specifically look at the percentage of high confidence estimates to the others

```{r confperc, echo=FALSE,message=FALSE,warning=FALSE}
# Create a confidence variable
tmp$confidence<-0L; tmp$confidence[tmp$paired==4 | tmp$paired==0] <- 1L
tmp %<>% mutate(confidence=as.factor(confidence))

ploty<-tmp%>%
  group_by(targ_db, confidence) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count))%>%
  filter(confidence==1)%>%
  ggplot()+geom_bar(aes(targ_db,percentage),fill="purple",stat="identity")+
  xlab("Database")+ylab("Percentage of Confident Pairings")+
  scale_y_continuous(labels = scales::label_number(accuracy = 0.01)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plotly::ggplotly(ploty)%>% plotly::layout(yaxis = list(hoverformat = '.4f'))
```

<br><br>

How does the confidence in the pairing vary per hazard?

```{r hazperc, echo=FALSE,message=FALSE,warning=FALSE}

ploty<-tmp2<-do.call(rbind,lapply(hazzies,function(haz){
  tmp%>%filter(grepl(haz,targ_hzAb,ignore.case = F))%>%
    group_by(confidence)%>%
    summarise(count = n()) %>%
  mutate(percentage = count / sum(count),
         Hazard=haz)%>%
  filter(confidence==1)%>%dplyr::select(Hazard,percentage)
}))%>%filter(Hazard%in%manyhaz)%>%
  ggplot()+geom_bar(aes(Hazard,percentage),fill="forestgreen",stat="identity")+
  xlab("Hazard")+ylab("Percentage of Confident Pairings")+
  scale_y_continuous(labels = scales::label_number(accuracy = 0.01))
  
rm(tmp,tmp2)

plotly::ggplotly(ploty)%>% plotly::layout(yaxis = list(hoverformat = '.4f'))
```

## Modelling

Let's look at the feature importance (FIRM-ICE method):

```{r ftimport, echo=FALSE,message=FALSE,warning=FALSE}
# Read it in
ftimp<-readRDS("../../Analysis_Results/Pairing/FeatureImportance.RData")
# Show a table of everything except the database variables
ftimp[!grepl("`",ftimp$Variable),]%>%filter(!grepl("-",ftimp$Variable))%>%
  mutate(Directionality=case_when(Directionality==-1 ~ "Negative",
                                  Directionality==1 ~ "Positive"))%>%
  kable(align="l", format = "html", digits = c(0, 1), col.names = c("Variable","Importance [%]","Directionality"))%>%
  kable_styling(c("bordered", "condensed"), full_width = T)

```

<br><br>

How likely was each database to be paired?

```{r dbimport, echo=FALSE,message=FALSE,warning=FALSE}
# Show a table of everything except the database variables
ftimp[!grepl("`",ftimp$Variable),]%>%filter(grepl("-",ftimp$Variable))%>%
  mutate(Importance=Importance*Directionality,
         Directionality=case_when(Directionality==1~"Likely",
                                  Directionality==-1~"Unlikely"))%>%
  dplyr::select(Variable,Directionality,Importance)%>%
  kable(align="l", format = "html", digits = c(0, 0, 1), col.names = c("Variable","Pairing Probability","Importance [%]"))%>%
  kable_styling(c("bordered", "condensed"), full_width = T)

```


































