---
title: "Monty Metadata Documentation"
author: "Hamish Patten"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(magrittr)
library(tidyverse)

# Function to extract all of the variable names directly from the Monty JSON schema
MontyJSONdocumentation<-function(){
  # JSON schema
  schemy<-jsonlite::fromJSON("../../Taxonomies/Montandon_Schema_V1-00.json")
  # Unravelled JSON scheme
  unjson<-plyr::llply(schemy, unlist)
  
  #@@@@@@@@@@@@@@@@@@@@@@@ Properties Field @@@@@@@@@@@@@@@@@@@@@@@#
  # First extract the names from the taxonomies section of the schema
  jsy<-names(unjson$properties)
  # Find titles, descriptions and data type
  indies<-which(grepl("taxonomies",jsy) &!grepl("enum",jsy) & !grepl("description",jsy) &
    !grepl("type",jsy) & !grepl("codelist",jsy) & !grepl("openCodelist",jsy) &
    !grepl("\\$ref",jsy) & !grepl("minItems",jsy) & !grepl("uniqueItems",jsy) & !grepl(".format",jsy))
  # but we want to make sure to add descriptions and data types to this
  indies<-sort(c(indies,indies+1,indies+2))
  # clean it up
  taxy<-as.data.frame(unjson$properties[indies]); colnames(taxy)<-"var_desc_name"
  # Add the important variables about this variable: coded name and the location in the data model/JSON schema
  taxy%<>%cbind(do.call(rbind,lapply(rownames(taxy),function(x) {
    y<-as.character(str_split(x,pattern = "\\.",simplify = T))
    data.frame(coded_name=y[(length(y)-1)],
               type=y[length(y)],
               field_layer=paste0(y[1:(length(y)-2)],collapse = "/"))
  })))
  # JSON objects come up as type="$ref" so we need to change this
  taxy$var_desc_name[taxy$type=="$ref"]<-"nested-object"; taxy$type[taxy$type=="$ref"]<-"type"
  taxy$type[taxy$type=="format"]<-"type"
  # Now convert this to wide format to correctly display the information
  taxy<-cbind(filter(taxy,type=="title")%>%dplyr::select(-type),
              filter(taxy,type=="description")%>%dplyr::select(1)%>%setNames("description"),
              filter(taxy,type=="type")%>%dplyr::select(1)%>%setNames("var_type"))
  # Cleanup
  rownames(taxy)<-NULL; taxy<-taxy[,c(2,1,3:ncol(taxy))]
  
  #@@@@@@@@@@@@@@@@@@@@@@@ Definitions Field @@@@@@@@@@@@@@@@@@@@@@@#
  jsy<-names(unjson$`$defs`)
  # Find titles, descriptions and data type
  indies<-which(grepl("title",jsy) & !grepl("enum",jsy) & !grepl("description",jsy) & !grepl("type",jsy) & 
                  !grepl("codelist",jsy) & !grepl("openCodelist",jsy) & !grepl("\\$ref",jsy) & 
                  !grepl("minItems",jsy) & !grepl("uniqueItems",jsy) & !grepl(".format",jsy) &
                  !grepl("pattern",jsy) & !grepl("required",jsy) & !grepl("minLength",jsy))
  # but we want to make sure to add descriptions and data types to this
  indies<-sort(c(indies,indies+1,indies+2))
  # clean it up
  defy<-as.data.frame(unjson$`$defs`[indies]); colnames(defy)<-"var_desc_name"
  # Add the important variables about this variable: coded name and the location in the data model/JSON schema
  defy%<>%cbind(do.call(rbind,lapply(rownames(defy),function(x) {
    y<-as.character(str_split(x,pattern = "\\.",simplify = T))
    data.frame(coded_name=y[(length(y)-1)],
               type=y[length(y)],
               field_layer=paste0(y[1:(length(y)-2)],collapse = "/"))
  })))
  # JSON objects come up as type="$ref" so we need to change this
  defy$var_desc_name[defy$type=="$ref"]<-"nested-object"; defy$type[defy$type=="$ref"]<-"type"
  defy$type[defy$type=="format"]<-"description"
  # Now convert this to wide format to correctly display the information
  defy<-cbind(filter(defy,type=="title")%>%dplyr::select(-type),
              filter(defy,type=="description")%>%dplyr::select(1)%>%setNames("description"),
              filter(defy,type=="type")%>%dplyr::select(1)%>%setNames("var_type"))
  # Cleanup
  rownames(defy)<-NULL; defy<-defy[,c(2,1,3:ncol(defy))]
  
  #@@@@@@@@@@@@@@@@@@@@@@@ Combined @@@@@@@@@@@@@@@@@@@@@@@#
  return(rbind(taxy,defy)%>%arrange(field_layer))
}
# Extract this
metdoc<-MontyJSONdocumentation()%>%filter(!stringi::stri_endswith(field_layer,fixed = "_obj"))

```

## Monty Metadata Documentation

This page is to help users of the Monty - Global Crisis Data Bank to understand the metadata of the database. This documentation therefore provides a detailed explanation of the data model of Monty. In combination with the Monty API, which can be found [here](montandon.westeurope.cloudapp.azure.com:8000/__docs__/), users of Monty data should have a more complete picture of the unique structure of the database.

```{r json_doc, echo=FALSE}
 metdoc %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling()
```


























